### 2021-07-13 paper/paper.tex
vpn edits to skater paper

\documentclass{bioinfo}
\copyrightyear{2016} \pubyear{2016}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing

% hyperref makes the margins screwy.
% https://groups.google.com/forum/#!topic/latexusersgroup/4W_SwGk6zx4
% http://ansuz.sooke.bc.ca/software/latex-tricks.php
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}

\access{Advance Access Publication Date:   }
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Genetics and Population Analysis}

\title[skater package for kinship analysis]{skater: An R package for SNP-based Kinship Analysis, Testing, and
Evaluation}

\author[FirstAuthorLastName \textit{et~al}.]{
Stephen D. Turner\,\textsuperscript{1},
V.P. Nagraj\,\textsuperscript{1},
Matthew Scholz\,\textsuperscript{1},
Shakeel Jessa\,\textsuperscript{1},
Carlos Acevedo\,\textsuperscript{1},
Jianye Ge\,\textsuperscript{2},
August Woener\,\textsuperscript{2},
Bruce Budowle\,\textsuperscript{2},
}

\address{
\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia, 22911, USA.\\
\textsuperscript{2}University of North Texas Health Science Center, Fort Worth, Texas,
76107, USA.\\
}

\corresp{To whom correspondence should be addressed. E-mail:
sturner@signaturescience.com}

\history{Received on XXX; revised on XXX; accepted on XXX}

\editor{Associate Editor: XXX}

\abstract{
\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and
IBD segment analysis methods produces results that often require further
downstream processing and manipulation. A dedicated software package
that consistently and intuitively implements this analysis functionality
is needed.\\
\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based
\textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and
\textbf{e}valuation with \textbf{R}. The skater package contains a suite
of well-documented tools for importing, parsing, and analyzing pedigree
data, performing relationship degree inference, benchmarking
relationship degree classification, and summarizing IBD segment data.\\
\textbf{Availability:} The skater package is implemented as an R package and is released under
the MIT license at https://github.com/signaturescience/skater.
Documentation is available at https://signaturescience.github.io/skater.\\
\textbf{Contact:}sturner@signaturescience.com\\
\textbf{Supplementary information:} Supplementary data are available at Bioinformatics Online.}

\maketitle

\section{Introduction}

Inferring familial relationships between individuals using genetic data
is a common problem in population genetics, medical genetics, and
forensics. There are multiple approaches to estimating relatedness
between samples, including genome-wide measures such as those
implemented in Plink \citep{purcell2007} or KING
\citep{manichaikul2010}, and methods that rely on identity by descent
(IBD) segment detection such as GERMLINE \citep{gusev2009}, hap-IBD
\citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts focusing
on benchmarking these methods \citep{ramstetter2017, vries2021} have
been aided by tools for simulating pedigrees and genome-wide SNP data
\citep{caballero2019}. Analyzing results from genome-wide SNP-based
kinship analysis or comparing analyses to simulated data for
benchmarking have to this point required writing one-off analysis
functions or utility scripts that are seldom distributed with robust
documentation, test suites, or narrative examples of usage. There is a
need in the field for a well-documented software package with a
consistent design and API that contains functions to assist with
downstream manipulation, benchmarking, and analysis of SNP-based kinship
assessment methods. Here we present the skater package for
\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
and \textbf{e}valuation with \textbf{R}.

\section{The skater package}

The skater package provides an intuitive collection of analysis and
utility functions for SNP-based kinship analysis. Functions in the
package include tools for importing, parsing, and analyzing pedigree
data, performing relationship degree inference, benchmarking
relationship degree classification, and summarizing IBD segment data.
The package is designed to adhere to ``tidy'' data analysis principles,
and builds upon the tools released under the tidyverse R ecosystem
\citep{Wickham2019}.

\subsection{Pedigree parsing, manipulation, and analysis}

The skater package has several functions for importing, parsing, and
analyzing pedigree data. Pedigrees define familial relationships in a
hierarchical structure. Many genomics tools for working with pedigrees
start with a .fam file, which is a tabular format with one row per
individual and columns for unique IDs of the mother, father, and the
family unit. The skater package contains the function
\texttt{read\_fam()} to read in a PLINK-formatted .fam file, and another
function \texttt{fam2ped()} to convert the content into a pedigree
object as a nested tibble with one row per family. All pedigree
processing from skater internally leverages a data structure from the
kinship2 package \citep{kinship2}. Further functions such as
\texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of
the pedigree for each family, while \texttt{ped2kinpair()} produces a
pairwise list of relationships between all individuals in the data with
the expected kinship coefficients for each pair.

\subsection{Relationship degree inference and benchmarking}

The skater package includes functions to translate kinship coefficients
to relationship degrees. The kinship coefficients could come from
\texttt{ped2kinpair()} or other kinship estimation software.

The \texttt{dibble()} function creates a \textbf{d}egree
\textbf{i}nference \textbf{t}ibble, with degrees up to the specified
maximum degree resolution, expected kinship coefficient, and lower and
upper inference ranges as defined in \citet{manichaikul2010}. The
\texttt{kin2degree()} function infers the relationship degree given a
kinship coefficient and a maximum degree resolution (e.g., 7th-degree
relatives) up to which anything more distant is treated as unrelated.

Once estimated kinship is converted to degree, it may be of interest to
compare the inferred degree to truth. When aggregated over many
relationships and inferences, this can help benchmark performance of a
particular kinship analysis method. The skater package adapts a
\texttt{confusion\_matrix()} function from \citet{clark2021} to provide
standard contingency table metrics (e.g.~sensitivity, specificity, PPV,
precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
The R-RMSE metric is defined more thoroughly in the skater package
vignette (see Supplementary Material), and may be a preferrable measure
of classification accuracy when benchmarking relationship degree
estimation. Taking the reciprocal of the target and predicted degree in
a typical RMSE calculation results in larger penalties for more
egregious misclassifications (e.g., classifying a first-degree relative
pair as second degree) than misclassifications at more distant
relationships (e.g., classifying a fourth-degree relative pair as
fifth-degree).

\subsection{IBD segment analysis}

Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020}
detect shared IBD segments between individuals. The skater package
includes functionality to take those IBD segments, compute shared
genomic centimorgan (cM) length, and convert that shared cM to a kinship
coefficient. In addition to inferred segments, these functions can
estimate ``truth'' kinship from simulated IBD segments
\citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise
IBD segments from IBD inference tools and from simulated IBD segments.
The \texttt{read\_map()} function reads in genetic map in a standard
format which is required to translate the total centimorgans shared IBD
to a kinship coefficient using the \texttt{ibd2kin()} function.

\section{Conclusion}

The skater R package provides a robust software package for data import,
manipulation, and analysis tasks typically encountered when working with
SNP-based kinship analysis tools. All package functions are internally
documented with examples, and the package contains a vignette
demonstrating usage, inputs, outputs, and interpretation of all key
functions (see Supplementary Material). The package contains internal
tests that are automatically run with continuous integration via GitHub
Actions whenever the package code is updated. The skater package is
permissively licensed (MIT) and is easily extensible to accommodate
outputs from new genome-wide relatedness and IBD segment methods as they
become available.

\section*{Funding}
\addcontentsline{toc}{section}{Funding}

\textbf{UNT FOLKS: NEED YOUR INPUT HERE!} The following rules should be
followed:

\begin{itemize}
\tightlist
\item
  The sentence should begin: `This work was supported by \ldots{}' -
\item
  The full official funding agency name should be given, i.e.~`National
  Institutes of Health', not `NIH' (full RIN-approved list of UK funding
  agencies)
\item
  Grant numbers should be given in brackets as follows: `{[}grant number
  xxxx{]}'
\item
  Multiple grant numbers should be separated by a comma as follows:
  `{[}grant numbers xxxx, yyyy{]}'
\item
  Agencies should be separated by a semi-colon (plus `and' before the
  last funding agency)
\item
  Where individuals need to be specified for certain sources of funding
  the following text should be added after the relevant agency or grant
  number `to {[}author initials{]}'.
\end{itemize}

An example is given here: `This work was supported by the National
Institutes of Health {[}AA123456 to C.S., BB765432 to M.H.{]}; and the
Alcohol \& Education Research Council {[}hfygr667789{]}.'

Oxford Journals will deposit all NIH-funded articles in PubMed Central.
See Depositing articles in repositories -- information for authors for
details. Authors must ensure that manuscripts are clearly indicated as
NIH-funded using the guidelines above.

% Bibliography
\bibliographystyle{natbib}
\bibliography{bibliography.bib}

\end{document}