### 2021-07-20 README.Rmd
link to documentation in readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# skater

<!-- badges: start -->
[![R-CMD-check-stable](https://github.com/signaturescience/skater/workflows/R-CMD-check-stable/badge.svg)](https://github.com/signaturescience/skater/actions)

[![R-CMD-check-dev](https://github.com/signaturescience/skater/workflows/R-CMD-check-dev/badge.svg)](https://github.com/signaturescience/skater/actions)
<!-- badges: end -->

**S**NP-based **K**inship **A**nalysis **T**esting and **E**valuation: miscellaneous **R** data analysis utilties.

## Installation

Install from GitHub:

``` r
remotes::install_github("signaturescience/skater", build_vignettes=TRUE)
```

<!-- A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT` or supplied as a string to `auth_token`. See `?remotes::install_github`.  -->

## Usage

The ["Basic Usage" vignette](https://signaturescience.github.io/skater/articles/basic_usage.html) steps through the primary functionality of the `skater` package:

``` r
vignette("basic_usage", package = "skater")
```

Full documentation: <https://signaturescience.github.io/skater/>.