### 2021-11-29 paper/paper.Rmd
first cut at moving to F1000 research

---
title: "**skater**: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
author: 
  - name: Stephen D. Turner
    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
  - name: V. P. Nagraj
    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
  - name: Matthew Scholz
    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
  - name: Shakeel Jessa
    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
  - name: Carlos Acevedo
    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
  - name: Jianye Ge
    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
  - name: August E. Woerner
    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
  - name: Bruce Budowle
    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
# Abstract up to 300 words. No citations, minimal abbreviations.
abstract: | 
    | 
    | **Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation. A dedicated software package that consistently and intuitively implements this analysis functionality is needed.  
    | **Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**. The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.  
    | **Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater. Documentation is available at https\:\/\/signaturescience.github.io/skater.
keywords: bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness
bibliography: bibliography.bib
output: BiocWorkflowTools::f1000_article
---

**R version**: `r R.version.string`

<!-- **Bioconductor version**: `r BiocManager::version()` -->

<!--  Update the name of the package below to be the name of the workflow package you are working on  -->

**skater package version**: `r packageVersion("skater")`

----

# Introduction

Inferring familial relationships between individuals using genetic data is a common practice in population genetics, medical genetics, and forensics. There are multiple approaches to estimating relatedness between samples, including genome-wide measures, such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection, such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020]. Recent efforts focusing on benchmarking these methods [@ramstetter2017] have been aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019]. Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are seldom distributed with robust documentation, test suites, or narrative examples of usage. There is a need in the field for a well-documented software package with a consistent design and API that contains functions to assist with downstream manipulation, benchmarking, and analysis of SNP-based kinship assessment methods. Here we present the skater package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.

# Methods

## Implementation

<!-- For software tool papers, this section should address how the tool works and any relevant technical details required for implementation of the tool by other developers. -->

The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis. Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data, described in full in the _Use Cases_ section below. The package adheres to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].

The skater package is hosted in the Comprehensive R Archive Network (CRAN) which is the main repository for R packages: <http://CRAN.R-project.org/package=skater>. Users can install skater in R by executing the following code:

```{r, eval=FALSE}
install.packages("skater")
```

Alternatively, the development version of skater is available on GitHub at <https://github.com/signaturescience/skater>. The development version may contain new features which are not yet available in the version hosted on CRAN. This version can be installed using the `install_github()` function in the devtools package:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("signaturescience/skater", build_vignettes=TRUE)
```

When installing skater, other packages which skater depends on are automatically installed, including magritr, tibble, dplyr, tidyr, readr, purrr, kinship2, corrr, rlang, and others.

## Operation

<!-- This part of the methods should include the minimal system requirements needed to run the software and an overview of the workflow for the tool for users of the tool. -->

Minimal system requirements for installing and using skater include R (version 3.0.0 or higher) and several tidyverse packages [@Wickham2019] that many R users will already have installed. Use cases are demonstrated in detail below. In summary, the skater package has functions for:

- Reading in various output files produced by commonly used tools in SNP-based kinship analysis
- Pedigree parsing, manpulation, and analysis
- Relationship degree inference
- Benchmarking and assessing relationship classification accuracy
- IBD segment analysis post-processing

A comprehensive reference for all the functions in the skater package is available at <https://signaturescience.github.io/skater/>.

<!-- # Results  -->

<!-- Optional - only if novel data or analyses are included -->

<!-- This section is only required if the paper includes novel data or analyses, and should be written as a traditional results section. -->

# Use Cases

<!-- Optional - only if NO new datasets are included -->

<!-- This section is required if the paper does not include novel data or analyses.  -->
<!-- Examples of input and output files should be provided with some explanatory context.  Any novel or complex variable parameters should also be explained in sufficient detail to allow users to understand and use the tool's functionality. -->

## Pedigree parsing, manipulation, and analysis

The skater package has several functions for importing, parsing, and analyzing pedigree data. Pedigrees define familial relationships in a hierarchical structure. Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit. The skater package contains the function `read_fam()` to read in a PLINK-formatted .fam file and another function `fam2ped()` to convert the content into a pedigree object as a nested tibble with one row per family. All pedigree processing from skater internally leverages a data structure from the kinship2 package [@sinnwell2014]. Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).

## Relationship degree inference and benchmarking

The skater package includes functions to translate kinship coefficients to relationship degrees. The kinship coefficients could come from `ped2kinpair()` or other kinship estimation software.

The `dibble()` function creates a **d**egree **i**nference t**ibble**, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010. The `kin2degree()` function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness. When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method. The skater package adapts a `confusion_matrix()` function from @clark2021 to provide standard contingency table metrics (e.g. sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more thoroughly in the skater package vignette and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation. In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number. Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).

## IBD segment analysis

Tools such as hap-IBD [@zhou2020], and IBIS [@seidman2020] detect shared IBD segments between individuals. The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and converts that shared cM to a kinship coefficient. In addition to inferred segments, these functions can estimate "truth" kinship from simulated IBD segments [@caballero2019]. The `read_ibd()` function reads pairwise IBD segments from IBD inference tools and from simulated IBD segments. The `read_map()` function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the `ibd2kin()` function.

<!-- # Discussion  -->

<!-- Optional - only if novel data or analyses are included -->

<!-- This section is only required if the paper includes novel data or analyses, and should be written in the same style as a traditional discussion section. -->

<!-- Please include a brief discussion of allowances made (if any) for controlling bias or unwanted sources of variability, and the limitations of any novel datasets. -->

<!-- # Conclusions  -->

<!-- Optional - only if novel data or analyses are included -->

<!-- This section is only required if the paper includes novel data or analyses, and should be written as a traditional conclusion. -->

# Summary

<!-- Optional - only if NO new datasets are included -->

The skater R package provides a robust software package for data import, manipulation, and analysis tasks typically encountered when working with SNP-based kinship analysis tools. All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions. The package contains internal tests that are automatically run with continuous integration via GitHub Actions whenever the package code is updated. The skater package is permissively licensed (MIT) and is easily extensible to accommodate outputs from new genome-wide relatedness and IBD segment methods as they become available.

<!-- # Data availability  -->

<!-- Optional - only if novel data or analyses are included -->

<!-- Please add details of where any datasets that are mentioned in the paper, and that have not have not previously been formally published, can be found.  If previously published datasets are mentioned, these should be cited in the references, as per usual scholarly conventions. -->

# Software availability

<!-- This section will be generated by the Editorial Office before publication. Authors are asked to provide some initial information to assist the Editorial Office, as detailed below. -->

<!-- 1. URL link to where the software can be downloaded from or used by a non-coder (AUTHOR TO PROVIDE; optional) -->
<!-- 2. URL link to the author's version control system repository containing the source code (AUTHOR TO PROVIDE; required) -->
<!-- 3. Link to source code as at time of publication (*F1000Research* TO GENERATE) -->
<!-- 4. Link to archived source code as at time of publication (*F1000Research* TO GENERATE) -->
<!-- 5. Software license (AUTHOR TO PROVIDE; required) -->

1. Software available from: <http://CRAN.R-project.org/package=skater>.
1. Source code available from: <https://github.com/signaturescience/skater>.
1. Archived source code at time of publication: **FIXME Zenodo DOI to come here**.
1. Software license: MIT License.

# Author information

SDT and VPN developed the R package.

All authors contributed to method development.

SDT wrote the first draft of the manuscript.

All authors assisted with manuscript revision.

All authors read and approved the final manuscript.

# Competing interests

<!-- All financial, personal, or professional competing interests for any of the authors that could be construed to unduly influence the content of the article must be disclosed and will be displayed alongside the article. If there are no relevant competing interests to declare, please add the following: 'No competing interests were disclosed'. -->

No competing interests were disclosed.

# Grant information

<!-- Please state who funded the work discussed in this article, whether it is your employer, a grant funder etc. Please do not list funding that you have that is not relevant to this specific piece of research. For each funder, please state the funder's name, the grant number where applicable, and the individual to whom the grant was assigned. If your work was not funded by any grants, please include the line: 'The author(s) declared that no grants were involved in supporting this work.' -->

This work was supported in part by award 2019-DU-BX-0046 (Dense DNA Data for Enhanced Missing Persons Identification) to B.B., awarded by the National Institute of Justice, Office of Justice Programs, U.S. Department of Justice and by internal funds from the Center for Human Identification. The opinions, findings, and conclusions or recommendations expressed are those of the authors and do not necessarily reflect those of the U.S. Department of Justice.

<!-- # Acknowledgments -->

<!-- This section should acknowledge anyone who contributed to the research or the article but who does not qualify as an author based on the criteria provided earlier (e.g. someone or an organization that provided writing assistance). Please state how they contributed; authors should obtain permission to acknowledge from all those mentioned in the Acknowledgments section. -->

<!-- Please do not list grant funding in this section. -->

<!--

# USING R MARKDOWN

Some examples of commonly used markdown syntax are listed below, to help you get started.

## Cross-references

For portability between different output formats, use the syntax introduced by *bookdown*, such as `(\#label)` for labels and `\@ref(label)` for cross-references. The following sections provide examples of referencing tables, figures, and equations.

## Citations

You can include references in a standard Bibtex file.  The name of this file is given in the header of the markdown document (in our case it is *sample.bib*).  References to entries in the Bibtex file are made using square brackets and use an @ plus the key for the entry you are referencing [@Smith:2012qr].  You can combine multiple entries by separating them with a semi-colon [@Smith:2012qr; @Smith:2013jd].
The default bibliography style uses numerical citations.  For superscript or author-year citations set the header metadata field `natbiboptions` to either `super` or `round`, respectively.

If you specify a figure caption to a code chunk using the chunk option `fig.cap`, the plot will be automatically labeled and numbered. The figure label is generated from the label of the code chunk by prefixing it with `fig:`, e.g., see Figure \@ref(fig:plot).

## Tables

Markdown syntax tends to lack some of the more sophisticated formatting features available in LaTeX, so you may need to edit the tables later to get the desired format.

| First name  | Last Name | Grade |
| ----------- | --------- | ----- |
| John        | Doe       |   7.5 |
| Richard     | Miles     |     2 |

Table: Caption to table.

Just like figures, tables with captions will also be numbered and can be referenced. Captions are entered as a paragraph beginning with the string "Table:" (or just ":"), which may appear either before or after the table. A label for the table should appear in the beginning of the caption in the form of `(\#tab:label)`, e.g., see Table \@ref(tab:table).

: (\#tab:table) A table with text justification.

| First name  | Last Name | Grade |
| ----------- | :-------: | ----: |
| John        | Doe       |   7.5 |
| Richard     | Miles     |     2 |

## Figures

You can include static figures (i.e. no generated by code) using the `include_graphics()` function from the **knitr** package, in a standard code chunk.

You can again use the `fig.cap` option to provide the figure caption, and reference the image based on the code chunk label.  You can also use options such as `fig.align` and `fig.width` to adjust the position and size of the image within the final document, e.g. Figure \@ref(fig:frog-picture) is a frog.

Alternatively, you can use the standard markdown syntax like so:

![This is a smaller version of the same picture, inserted using the standard markdown syntax](frog.jpg){width=2cm height=2cm}

Please give figures appropriate filenames, e.g.: figure1.pdf, figure2.png.

Figure legends should briefly describe the key messages of the figure such that the figure can stand alone from the main text. However, all figures should also be discussed in the article text. Each legend should have a concise title of no more than 15 words. The legend itself should be succinct, while still explaining all symbols and abbreviations. Avoid lengthy descriptions of methods.

For any figures reproduced from another publication (as long as appropriate permission has been obtained from the copyright holder -see under the heading 'Submission'), please include a line in the legend to state that: 'This figure has been reproduced with kind permission from [include original publication citation]'.

-->